Comment: Inferring Trade Costs from Trade Booms and Trade Busts

### 2016-06-03 Rédaction/Manuscript.tex
Redémarrage ?

\cite{JMN2011} suggest\footnote{The method is also developed in
\cite{JMN2008}, \cite{JMN2010}, \cite{NOVY}.} a method for
inferring international trade costs from trade flows. They
define trade costs as all barriers to trade, notably transport
and transaction costs. Using the general equilibrium model of
\cite{AW2003}, the authors calculate those costs and their
evolution during the first and second waves of globalization
(1870-1913 and 1950-2000) as well as the interwar period
(1921-1939) thanks to the impressive set of data they collected
on trade flows, GDP and exports between 27
countries\footnote{Argentina, Australia, Austria, Belgium,
Brazil, Canada, Denmark, France, Germany, Greece, India,
Indonesia, Italy, Japan, Mexico, the Netherlands, New Zealand,
Norway, the Philippines, Portugal, Spain, Sri Lanka, Sweden,
Switzerland, the United Kingdom, the United States, and
Uruguay. The data contain 130 country pairs.}. They provide a
decomposition of the growth of trade caused by the reduction in
trade costs and economic expansion. Their main result concerns
a difference of nature between the two globalizations:

@@

We argue that this result is actually driven by the ad hoc
method of aggregation. When relying instead on a
three-countries version of \cite{AW2003}'s model for
aggregating bilateral trade costs - where the three countries
are the country of interest, its aggregate trade partners
present in the data, and the rest of the world - the alleged
difference of natures between the two globalizations fades
away. We show that the authors' main results is not robust to
the aggregation technique because the one they use also
captures structure effects. The inferred difference between the
two globalization concerns the size of the countries comes from
the phenomenon of globalization, not trade costs. Moreover, we
cast doubt on the possibility to reach conclusions of this
nature through an approach that relies on the study of trade
flows only. Deducing trade costs from trade flows, and using
trade costs to explain trade flows is essentially a circular
reasoning. Therefore, Jacks et al.'s approach cannot be an
alternative to traditional investigations of impediments to
trade, such as commodity price gaps.

This paper starts by presenting Jacks et al.'s approach to the
measure of trade costs, and insists on its relevance at the
bilateral level to control for multilateral trade barriers in
gravity regressions. It then highlights how paradoxical the
result on a difference of nature between the two globalizations
is since it cannot be deducted from a comparison of the
evolution of openness ratios. Section \ref{2} shows that the
conclusion is only driven by the authors' ad hoc aggregation
method. We propose a microfounded way to aggregate trade costs
and the puzzle fades away. Section \ref{3} explores the reasons
why Jacks et al.'s aggregation technique ends up providing
different results for the two globalizations. We argue that
what Jacks et al. misleadingly attribute to unequal trade costs
decreases between the two globalizations is instead a
difference in the distribution of trade growth over trading
partners.

@@

Although it is consistent with many models of international
trade, Jacks, Meissner and Novy's work is primarily based on
the general equilibrium model framework of \cite{AW2003}. $n$
countries, each represented by a maximizing consumer, exchange
goods over one single period. In this Armington world,
production is not modelized\footnote{In the working paper
version of \cite{JMN2010}, the authors provide a version of the
model with production. The key equation is identical to the one
of the demand-side model.} and each country is initially
endowed with a differentiated representative good. Trade occurs
because of consumers' taste for diversity. The preferences of
all countries are assumed to be identical and modelized by a
Constant Elasticity of Substitution (CES) utility function.

@@

Since the trade costs measure is assimilated to trade flows
relative to domestic activity, they should intuitively only
reformulate an information already contained in any measure of
trade flows relative to economic size, like an openness ratio
defined for instance as the ratio of exportations to GDP.
However the main conclusion in Jacks et al.'s article, namely
that the first globalization was more driven by the decrease in
trade costs than the second one seems to suggest that the
circularity intuition is wrong. Decomposing the level of
exportations of a country as the product of its GDP and its
openness ratio yields the results presented\footnote{We follow
the authors in displaying the changes in the log of variables
in percentage. However, one should keep in mind that given the
size of the changes, they should not be taken as percentage
variations.} in table \ref{OR}. The increase in exportations of
all countries in the table is dominantly due to the growth of
GDP for the large majority of the 27 countries present in the
data. On average the decomposition attributes 74\% (183/246) of
the growth in trade to the increase in GDP in the second
globalization and 60\% in the first globalization. To the
contrary, Jacks et al. find that their measure of trade costs
explains about 60\% (107/178) of the growth in international
trade during the first globalization, the increase in GDP
explaining about\footnote{Contributions of growth in income
similarity and of change in multilateral factors are negative.}
45\%.

>>>>

Jacks and his coauthors offer in a number of papers \footnote{The method is developed in\cite{JMN2008}, \cite{JMN2010}, \cite{NOVY} and \cite{JMN2011}. We will use this latter paper as a referenc} a method for
inferring international trade costs from trade flows. Using the general equilibrium model of
\cite{AW2003}, they calculate trade costs (defined as as all barriers to trade, notably transport
and transaction costs) and their
evolution during the first and second waves of globalization
(1870-1913 and 1950-2000) as well as the interwar period
(1921-1939) thanks to the impressive set of data they collected
on trade flows, GDP and exports between 27
countries\footnote{Argentina, Australia, Austria, Belgium,
Brazil, Canada, Denmark, France, Germany, Greece, India,
Indonesia, Italy, Japan, Mexico, the Netherlands, New Zealand,
Norway, the Philippines, Portugal, Spain, Sri Lanka, Sweden,
Switzerland, the United Kingdom, the United States, and
Uruguay. The data contain 130 country pairs.}. They provide a
decomposition of the growth of trade caused by the reduction in
trade costs and economic expansion. They use their computations to underline
a difference of nature between the two globalizations:

@@

We argue that this result is actually driven by an ad hoc
method of aggregation. When relying instead on a
three-countries version of \cite{AW2003}'s model for
aggregating bilateral trade costs - where the three countries
are the country of interest, its aggregate trade partners
present in the data, and the rest of the world - the alleged
difference of natures between the two globalizations fades
away. \cite{JMN2011}'s aggregation technique captures structure effects. The inferred difference between the
two globalization
????? comes from the evolving size of countris rather that trade costs ????
concerns the size of the countries comes from
the phenomenon of globalization, not trade costs. More generally, we
cast doubt on the possibility to distinguish the effect of trade costs and economic expansion through an approach that relies only on the study of trade
flows. Deducing trade costs from trade flows, and then using
trade costs to explain trade flows is essentially a circular
reasoning. Therefore, Jacks et al.'s approach cannot be an
alternative to traditional investigations of impediments to
trade, such as commodity price gaps.

This paper starts by presenting Jacks et al.'s approach to the
measure of trade costs, and insists on its relevance at the
bilateral level to control for multilateral trade barriers in
gravity regressions (section \ref{1}). It then highlights how paradoxical the
result on a difference of nature between the two globalizations
is since it cannot be deducted from a comparison of the
evolution of openness ratios. Section \ref{2} shows that the
conclusion is only driven by the authors' ad hoc aggregation
method. We propose a microfounded way to aggregate trade costs
and the puzzle fades away. Section \ref{3} explores the reasons
why Jacks et al.'s aggregation technique ends up providing
different results for the two globalizations. We argue that
what Jacks et al. misleadingly attribute to unequal trade costs
decreases between the two globalizations is instead a
difference in the distribution of trade growth over trading
partners.

@@

Although it is consistent with many models of international
trade, Jacks, Meissner and Novy's work is primarily based on
the general equilibrium model framework of \cite{AW2003}. $n$
countries, each represented by a maximizing consumer, exchange
goods over one single period. In this Armington world,
production is not modelized\footnote{In the working paper
version of \cite{JMN2010}, the authors provide a version of the
model with production. The key equation is identical to the one
of the model without production.} and each country is initially
endowed with a differentiated representative good. Trade occurs
because of consumers' taste for diversity. The preferences of
all countries are assumed to be identical and modelized by a
Constant Elasticity of Substitution (CES) utility function.

@@

Since the trade costs measure is assimilated to trade flows
relative to domestic activity, they should intuitively only
reformulate an information already contained in any measure of
trade flows relative to economic size, like an openness ratio
defined for instance as the ratio of exportations to GDP.
However the main conclusion in Jacks et al.'s article, namely
that the first globalization was more driven by the decrease in
trade costs than the second one seems to suggest that the
circularity intuition is wrong. We decompose the level of
exportations of a country as the product of its GDP and its
openness ratio yields the results presented\footnote{We follow
the authors in displaying the changes in the log of variables
in percentage. However, one should keep in mind that given the
size of the changes, they should not be taken as percentage
variations.} in table \ref{OR}. The increase in exportations of
all countries in the table is dominantly due to the growth of
GDP for the large majority of the 27 countries present in the
data. On average the decomposition attributes 74\% (183/246) of
the growth in trade to the increase in GDP in the second
globalization and 60\% in the first globalization. To the
contrary, Jacks et al. find that their measure of trade costs
explains about 60\% (107/178) of the growth in international
trade during the first globalization, the increase in GDP
explaining about\footnote{Contributions of growth in income
similarity and of change in multilateral factors are negative.}
45\%.
